---
title: "New Cases Per Capita and Vaccination Ratios - by County Level"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: spacelab
---

```{r setup, include=FALSE}
#Import libraries
library(flexdashboard)
library(httr)
library(tidyverse)
library(jsonlite)
library(stargazer)
library(magrittr)
library(lmtest)
library(highcharter)
```

```{r}
#Import data, and manipulate

  
main <- "https://api.covidactnow.org/v2/"
state <- "counties.json?apiKey="
key <- "88d523069840437f851b3738a9cb1021"

api_covid <- paste0(main, state, key)

raw_covid <- GET(api_covid)

json_covid <- fromJSON(rawToChar(raw_covid$content), flatten = TRUE)

df <- json_covid

df <- df %>% select(state, 
                    county, 
                    fips, 
                    population, 
                    metrics.weeklyNewCasesPer100k, 
                    metrics.vaccinationsCompletedRatio,
                    metrics.vaccinationsAdditionalDoseRatio) %>% 
  rename(thirddose = metrics.vaccinationsAdditionalDoseRatio,
         seconddose = metrics.vaccinationsCompletedRatio,
         newcases = metrics.weeklyNewCasesPer100k) %>% 
  mutate(newcasedensity = newcases / population,
         seconddose2 = seconddose^2,
         thirddose2 = thirddose^2)

```

Figures and Regressions
===================================================================

Column {data-width = 450}
-----------------------------------------------------------------------
### Figure 1: Second-dose Completion Ratio and New Cases Per Capita
```{r}

ggplot(df, aes(seconddose, newcasedensity))+
  geom_point(alpha = 1/5, color = "blue")+
  ylim(0,0.05)+
  geom_smooth(color = "red", se = F)
```

### Figure 2: Third-dose Completion Ratio and New Cases Per Capita

```{r}

ggplot(df, aes(thirddose, newcasedensity))+
  geom_point(alpha = 1/5, color = "blue")+
  ylim(0,0.05)+
  geom_smooth(color = "red", se = F)
```

Column {data-width = 250}
------------------------------------------------------------------------

### About Figure 1

Several Observations have newcasedensity > 0.05. They are not shown in the
Figure 1 to improve informativity.

The fitted smooth line is a U-shaped curve. Therefore we add a quadratic term
of seconddose to detect whether the U-shaped relationship is significant. 



### About Figure 2

Similarly,

Several Observations have newcasedensity > 0.05. They are not shown in the
Figure 2 to improve informativity.

The fitted smooth line is a U-shaped curve. Therefore we add a quadratic term
of thirddose to detect whether the U-shaped relationship is significant. 




Column {data-width = 450}
-------------------------------------------------------------------------

### Regression 1 - Second-dose Ratio
```{r}

twodoses <- lm(newcasedensity ~ seconddose2 + seconddose, df)
summary(twodoses)
```

### Regression 2 - Third-dose Ratio
```{r}

threedoses <- lm(newcasedensity ~ thirddose2 + thirddose, df)
summary(threedoses)
```


Explanation
=================================================================

Our regression models indicate that with the growing second dose vaccination
completion ratio and third dose vaccination completion ratio, the new case density,
which is (new cases) / (population), drops at the beginning, but rises up after
reaching to a high level of the vaccination ratio (61.6% for second-dose, 34.3%
for third-dose). All coefficients are significant at 1% level.

Such a trend can be explained through population density ratio, policy changes,
and weaker policy implementation. Since our regressions are at the county level,
there are many counties with a low population density ratio. In these counties, 
social distance can be naturally and more easily achieved in public area, thus 
a lower new case density. Futhermore, counties with high vaccination ratios tend 
to loose the covid safety restrctions, which leads a higher new case density.
Similarly, residents in these counties also care relatively less if they strictly
obligate the safety restrictions, and expose themselves to a higher infection risk.

Thankfully, recent variations of coronavirus are not likely to cause severe symtoms
and death, so a sligtly higher new case density won't be a serious problem.




